Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS

### 2025-04-22 report/draft.qmd
update

Source: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.

@@

## Factor decomposition of wealth inequality

@@

## Group decomposition of wealth inequality by rural and urban
The following two sections present the group decomposition of wealth inequality by the urban-rural and regional divisions.

@@

Summary Statistics of Urban and Rural Income and Wealth, 2017 and 2019
:::

The summary statistics above shows that in both 2017 and 2019, the mean and median income levels are significantly higher in urban areas compared to rural areas. The mean wealth in urban areas is 3.48 times the mean wealth in rural areas in 2017, and 3.78 times in 2019. The median wealth in urban areas is 2.82 times the median wealth in rural areas in 2017, and 2.52 times in 2019.

Notably, the Gini coefficient for wealth is higher in rural areas than in urban areas in both 2017 and 2019. The Gini coefficient for both urban rural areas declined from 2017 to 2019.

The decomposition of the Gini coefficient into within-group, between-group, and residual components shows[@lambert_inequality_1993]:

@@

@@

Decomposition of Wealth Gini Coefficient by Urban-Rural Division, 2017 and 2019 (%)
:::

This decomposition reveals that about 55-57% of total wealth inequality is due to inequality within urban and rural groups, while about 33-34% is due to inequality between urban and rural areas. The residual component, representing the interaction between the within and between components, accounts for about 10-12% of total inequality. The stability of these components between 2017 and 2019 suggests that the urban-rural parity remains a significant driver of wealth inequality in China.

## Group decomposition of wealth inequality by region

@@

## What's next
I will read more papers and contrast the results with data in this study. I am not sure about what else I can do for this paper. I would like to hear your suggestions. Thanks.

>>>>

All figures are reported in RMB at constant 2017 prices.

Sources: @tan_earnings_2017's calculation based on the 2011 CHFS data and authors' calculation based on the 2017 and 2019 CHFS data.

@@

## Factor decomposition of wealth inequality
This section examines the composition and distribution of household wealth in China by analyzing specific asset and debt categories using data from the 2017 and 2019 China Household Finance Survey (CHFS). This detailed decomposition provides insights into the relative importance of different components contributing to wealth inequality.

::: {#tbl-wealth-summary}

{{< include wealth_summary.txt >}}

All figures are reported in RMB at constant 2017 prices.

Summary Statistics of Wealth, 2017 and 2019
:::

The above table intuitively shows the state of household wealth allocation in China in 2017 and 2019. Clearly the mean and median value of house dominates other assets. It is also notable that despite the high inequality in housing assets, as it is shown in the next table, the proportion of households that have no houses at all is fairly low at just 10%. It is also surprising that the housing debt is very low compared to the asset value, and 82% of the households have no house debts.

@@

## Group decomposition of wealth inequality by rural and urban
Group decomposition of wealth inequality by rural and urban

To analyze the sources of wealth inequality, this paper employs the classic Gini decomposition identity initially developed by Bhattacharya and Mahalanobis (1967)[@bhattacharya_regional_1967]. The overall Gini coefficient  is decomposed as follows:

$$G = G_{B} + \sum_{k} a_{k}G_{k} + R$$

where $G_{B}$ is the between-group Gini, which captures inequality arising solely from differences in group means; $G_{k}$ is the within-group component, a weighted sum of the inequality within each group, with weights $a_{k}$ representing the products of each group's population share and wealth share; and $R$ is the overlap (residual) component, which captures inequality due to the intersection of group wealth distributions. Lambert and Aronson (1993) provide a geometric interpretation of this overlap term, demonstrating that $R$ corresponds to the Lorenz curve area generated by reranking when subgroup wealth distributions overlap. A positive $R$ thus indicates significant intersections among wealth distributions of different groups[@lambert_inequality_1993].

@@

Summary statistics of urban and rural income and wealth, 2017 & 2019
:::

The summary statistics highlight significant and persistent disparities between urban and rural China. Urban households consistently exhibit higher mean and median incomes and wealth compared to rural households. Urban mean wealth was 3.48 times greater than rural mean wealth in 2017 and widened slightly to 3.78 times in 2019. Median urban wealth was 2.82 times higher than the rural median in 2017, though this ratio decreased slightly to 2.52 in 2019. Additionally, the Gini coefficient for wealth is consistently higher in rural areas, though both rural and urban Gini coefficients declined modestly between 2017 and 2019.

@@

All figures are reported in RMB at constant 2017 prices.

@@

Decomposition of wealth Gini coefficient by urban–rural division, 2017 & 2019 (%)
:::

The decomposition results indicate that about 55–57% of total wealth inequality is attributable to within-group disparities among urban and rural households. The between-group component, reflecting the urban-rural wealth divide, contributes approximately 33–34%. The overlap component, representing inequality due to intersecting distributions of urban and rural wealth, accounts for around 9–12% of total inequality. The consistency of these shares between 2017 and 2019 underscores the continued significance of the urban-rural divide in shaping China's wealth inequality.

@@
